A Typed Programmatic Interface to Contracts on the Blockchain

### 2021-02-19 papers/execution-model/model.tex
major reorganization

%\begin{definition}%[Blockchain system]
A \emph{blockchain configuration} has the form
$\BLOCKCHAIN[ \NODE_1, \dots, \NODE_n]$, for some $n>0$, where $\BLOCKCHAIN$ is a
blockchain and the $\NODE_i$ are local nodes, for $1\le i\le n$.
Moreover, the accounts on the local nodes are all different and each
local account has a manager in $\BLOCKCHAIN$:
\begin{enumerate}
\item for all $1\le i< j\le n$, $\NODE_i.\ACCOUNTS \cap
  \NODE_j.\ACCOUNTS = \emptyset$;
\item for all $1\le i \le n$, 
    $\forall a \in \NODE_i.\ACCOUNTS \implies a.\PUK \in \DOM( \BLOCKCHAIN.\MANAGERS)$.
\end{enumerate}
%\end{definition}
\begin{definition}
  Preserved by a step on a well-formed configuration $ [\EXPRS, \ACCOUNTS] \|
  [\PENDING, \MANAGERS, \CONTRACTORS, \TIME]$:
  \todo[inline]{Is this preserved?}
  \begin{itemize}
  \item for all $\OPH$ in $\EXPRS$, $\OPH \in \DOM (\PENDING)$,
  \item for all $\PUK$ in $\EXPRS$, $\PUK \in \DOM (\MANAGERS)$,
  \item for all $\PUH$ in $\EXPRS$, $\PUH \in \DOM (\CONTRACTORS)$.
  \end{itemize}
\end{definition}

\clearpage{}
\subsection{Properties of blockchain state transitions}

If $[\PENDING, \MANAGERS, \CONTRACTORS, \TIME] \BlockTrans{}
[\PENDING', \MANAGERS', \CONTRACTORS', \TIME']$, then
\begin{itemize}
\item $\TIME \le \TIME'$
\item $\DOM (\PENDING) \subseteq \DOM(\PENDING')$
\item for all $\OPH \in \DOM(P)$, if 
  $\PENDING (\OPH) = \Angle{\OP,  \hat\TIME, \STATUS}$, then either
  \begin{itemize}
  \item     $\PENDING'    (\OPH) = \PENDING (\OPH)$; or
  \item $\STATUS  = \STATUSPENDING$ and $\PENDING' (\OPH) =
    \Angle{\OP, \hat\TIME, \STATUSTIMEOUT}$; or
  \item  $\STATUS  = \STATUSPENDING$, $\TIME - \hat\TIME
    \le 60 $, and
    $\PENDING' (\OPH) =       \Angle{\OP, \hat\TIME, \STATUSINCLUDED\
      \TIME}$.
  \end{itemize}
 \item for all $\OPH \in \DOM(P)$ and $\PENDING (\OPH) = \Angle{\OP,  \hat\TIME, \STATUSPENDING}$, 
    \begin{itemize}
    	\item  if $\MANAGERS(\OP.\PUK) =  \Angle{\BAL, \COU}$ then  
  $\COU.b = \TRUE$.
    \end{itemize}
 \item $ \DOM (\MANAGERS) \subseteq \DOM (\MANAGERS')$
 \item for all $\PUK \in \DOM (\MANAGERS)$\\
   if $\MANAGERS (\PUK) =
   \Angle{\BAL, \COU}$, 
   then $\MANAGERS' (\PUK) =
   \Angle{\BAL', \COU'}$ and
   \begin{itemize}
   \item if $\COU.b=\TRUE$ and $\COU'.b=\FALSE$, then $\COU.n' \in \{
     \COU.n,  \COU.n+1\}$,
   \item otherwise $\COU.n = \COU'.n$
   \item If $\COU.n = \COU'.n$, then $\BAL = \BAL'$.
   % \item if $\COU.b= \FALSE$, then $n' = n$
   % \item if $\COU.b= \TRUE$ and $b'=\TRUE$, then $n' =n$
   \end{itemize}

 \item $ \DOM (\CONTRACTORS) \subseteq \DOM (\CONTRACTORS')$
   \begin{itemize}
   \item for all $\PUH \in \DOM (\CONTRACTORS)$,
     $\CONTRACTORS (\PUH).\CODE = \CONTRACTORS' (\PUH).\CODE$
   \end{itemize}
 \item invariant for contractors:
   for all $\PUH \in \DOM (\CONTRACTORS)$,
   $\CONTRACTORS (\PUH) = \Angle{ \CODE, \TIME, \BAL, \STORAGE}$
   implies that $\PUH = \GENERATEHASH (\CODE, \TIME)$.
\end{itemize}

@@

To describe contracts we associate each public hash to a pair of
types, the parameter type and the storage type.
\begin{align*}
  \Delta & ::= \EmptyEnv \mid \PUH: \TPAIR\ \TYPE\ \TYPE, \Delta
\end{align*}

@@

\section{Transitions}
\label{sec:transitions}

We distinguish several kinds of transitions depending on their scope:
\begin{enumerate}
\item $\ExprTrans$ single-step evaluation of expressions in the node state,
\item $\NodeTrans$ internal transitions of a node,
\item $\BlockTrans$ transitions of the blockchain state,
\item $\SystemTrans$ blockchain system transitions.
\end{enumerate}

Evaluation of expressions is standard call-by-value lambda calculus defined
using evaluation contexts
$\EC{}$. Figure~\ref{fig:expression-reduction} shows some of the
reduction rules. 
\begin{figure}[tp]
  \begin{mathpar}
    \inferrule{}{ \EC{(\lambda x.\EXPR)\VAL} \ExprTrans
      \EC{\EXPR[\VAL/x]} }

    \inferrule{}{ \EC{\MATCH\ (\VAL_1, \VAL_2) \WITH\ (x_1,
        x_2)\to\EXPR \mid \dots} \ExprTrans \EC{\EXPR[\VAL_1,
        \VAL_2/x_1, x_2]} }

    \inferrule{}{ \EC{\MATCH\ \NIL\ \WITH\ \NIL\to\EXPR \mid \dots}
      \ExprTrans \EC{\EXPR} }

    \inferrule{}{ \EC{\MATCH\ (\CONS\ \VAL_1\ \VAL_2)\ \WITH\ (\CONS\
        x_1\ x_2)\to\EXPR \mid \dots} \ExprTrans
      \EC{\EXPR[\VAL_1,\VAL_2/x_1, x_2]} }

    \inferrule{}{ \EC{\MATCH\ \NONE\ \WITH\ \NONE\to\EXPR \mid \dots}
      \ExprTrans \EC{\EXPR} }

    \inferrule{}{ \EC{\MATCH\ (\SOME\ \VAL)\ \WITH\ (\SOME\ x)\to\EXPR
        \mid \dots} \ExprTrans \EC{\EXPR[\VAL/x]} }

    \inferrule{}{ \EC{\MATCH\ (\LEFT\ \VAL)\ \WITH\ (\LEFT\ x)\to\EXPR
        \mid \dots} \ExprTrans \EC{\EXPR[\VAL/x]} }

    \inferrule{}{ \EC{\MATCH\ (\RIGHT\ \VAL)\ \WITH\ (\RIGHT\
        x)\to\EXPR \mid \dots} \ExprTrans \EC{\EXPR[\VAL/x]} }

    \inferrule{\TYPE \SubType \TYPEU}{ \EC{\CAST\VAL\TYPE\TYPEU} \ExprTrans \EC{\VAL}}
  \end{mathpar}
  \caption{Select expression reduction rules}
  \label{fig:expression-reduction}
\end{figure}
The internal transitions of a node are just evaluation of expressions.
\begin{mathpar}
  \inferrule[Node-Eval]
  {
    \EXPR \ExprTrans \EXPR'
  }{
    [\EC\EXPR :: \EXPRS, \ACCOUNTS] \NodeTrans{}
    [\EC{\EXPR'} :: \EXPRS, \ACCOUNTS]
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[Node-Inject]{
    \Angle{\PAK,\PUK} \in \ACCOUNTS \\
    \CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ) \\
    \CHECKCOU (\MANAGERS, \PUK) \\
    \CHECKPUB (\MANAGERS, \PUK') \\
    \CHECKGAS (\PUK, \PUK', \NTEZ, \MTEZ) \\
    %\OPH = \GENERATEOPH (\PUK, \PUK','', \NTEZ, \MTEZ, \TIME) \\
    \OPH = \GENERATEOPH (\OP, \TIME) \\
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ    
  }{
    { [\EC\OP :: \EXPRS, \ACCOUNTS] \|
      [\PENDING, \MANAGERS, \CONTRACTORS, \TIME] } \SystemTrans \\
   { [\EC{\OPH}  :: \EXPRS, \ACCOUNTS] \|
     [ \OPH \mapsto \Angle{\OP, \TIME, \STATUSPENDING}
     ::\PENDING,
     \UPDATECOU(\MANAGERS, \PUK, \TRUE),
     \CONTRACTORS,
     \TIME]
   }
  }
\end{mathpar}

\begin{mathpar}
  \inferrule[Block-Accept]{
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ \\
    \TIME - \hat\TIME \le 60
  }{
    { 
      [\OPH \mapsto \Angle{\OP, \hat \TIME, \STATUSPENDING}
     ::\PENDING, \MANAGERS,
      \CONTRACTORS, \TIME]}
    \BlockTrans  \\
    { 
      [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSINCLUDING\ \TIME} :: \PENDING, 
      \UPDATESUCC(\MANAGERS, \PUK, \PUK', \NTEZ, \MTEZ),
      \CONTRACTORS, \TIME +1]}
  }
\end{mathpar}

The timeout transition applies to all operations.
\begin{mathpar}
  \inferrule[Block-Timeout]{
    % \OP = \TRANSFER[\STRING]\NTEZ\PUK\ADDR\MTEZ \\
    \TIME-\hat\TIME > 60
  }{ 
    {[\OPH \mapsto \Angle{\OP, \hat \TIME, \STATUSPENDING}
     ::\PENDING, \MANAGERS,
      \CONTRACTORS, \TIME]}
    \BlockTrans \\
    { 
      [\OPH \mapsto \Angle{\OP, \hat \TIME, \STATUSTIMEOUT}
     :: \PENDING, \UPDATECOU(\MANAGERS, \OP.\PUK, \FALSE),
      \CONTRACTORS, \TIME]}
  }
\end{mathpar}

\subsection{Special reductions}
\label{sec:special-reductions}

\begin{mathpar}
  \inferrule[Contract?]{
    \JTypeCode\CODE{\TPAIR\ \TYPE\ \TYPEU} \\
    \BLOCKCHAIN.\CONTRACTORS (\PUH) =  \Angle{\CODE, \tilde\TIME, \NTEZ', \STRING'}
    %
    }{
    { [\EC{\CONTRACT\ \TYPE\ \PUH}  :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    }
    \SystemTrans
    { [\EC{\SOME\ \PUH} :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
   }
  }

  \inferrule[Contract-Implicit?]{
    \PUK \in \DOM(\BLOCKCHAIN.\MANAGERS)
    %
    }{
    { [\EC{\CONTRACT\ \TUNIT\ \PUK}  :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    }
    \SystemTrans
    { [\EC{\SOME\ \PUK} :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
   }
  }

  \inferrule[Contract-No?]{
    \ADDR \in \DOM (\BLOCKCHAIN.\MANAGERS) \Rightarrow \TYPE\ne\TUNIT \\
    \BLOCKCHAIN.\CONTRACTORS (\ADDR) = \Angle{\CODE, \tilde\TIME,
      \NTEZ', \STRING'} \Rightarrow {}
        \JTypeCode\CODE{\TPAIR\ \TYPE'\ \TYPEU} \wedge \TYPE \ne \TYPE' \\
    }{
    { [\EC{\CONTRACT\ \TYPE\ \ADDR}  :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    }
    \SystemTrans
    { [\EC{\NONE} :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
   }
  }
\end{mathpar}

\subsection{Transitions on Nodes}
\label{sec:transitions-nodes}

Each node has (nondeterministic) rules to propose an operation.
When an operation $\OP$ appears, we check that the account is local by
looking up its public key $\OP.\PUK$ in the local accounts $\ACCOUNTS$ and
consider it signed with the corresponding private key $\PAK$.

\begin{mathpar}
  \inferrule[Node-System]
  {\NODE \NodeTrans \NODE'}
  { {\NODE :: \overline\NODE \| \BLOCKCHAIN}
    \SystemTrans
    {\NODE' :: \overline\NODE \| \BLOCKCHAIN}}

  \inferrule[Block-System]
  {\BLOCKCHAIN \BlockTrans \BLOCKCHAIN'}
  { {\overline\NODE \| \BLOCKCHAIN}
    \SystemTrans
    {\overline\NODE\| \BLOCKCHAIN'}}
\end{mathpar}

>>>>

%\begin{definition}%[Blockchain system]
A \emph{blockchain configuration} has the form
$\BLOCKCHAIN[ \NODE_1, \dots, \NODE_n]$, for some $n>0$, where $\BLOCKCHAIN$ is a
blockchain and the $\NODE_i$ are local nodes, for $1\le i\le n$.
In a \emph{well-formed configuration}, the accounts on the local nodes are all different and each
local account has a manager in $\BLOCKCHAIN$:
\begin{enumerate}
\item for all $1\le i< j\le n$, $\NODE_i.\ACCOUNTS \cap
  \NODE_j.\ACCOUNTS = \emptyset$;
\item for all $1\le i \le n$, 
    $\forall a \in \NODE_i.\ACCOUNTS \implies a.\PUK \in \DOM( \BLOCKCHAIN.\MANAGERS)$.
\end{enumerate}
%\end{definition}

\section{Operational Semantics}
\label{sec:transitions}

The operational semantics is defined by several kinds of transitions depending on their scope:
\begin{enumerate}
\item $\ExprTrans$ single-step evaluation of an expression in a local node,
\item $\NodeTrans$ internal transitions of a node,
\item $\BlockTrans$ transitions of the blockchain state,
\item $\SystemTrans$ blockchain system transitions.
\end{enumerate}

Evaluation of expressions is standard for call-by-value lambda calculus defined
using evaluation contexts
$\EC{}$. Figure~\ref{fig:expression-reduction} shows some of the
reduction rules. 
\begin{figure}[tp]
  \begin{mathpar}
    \inferrule{}{ \EC{(\lambda x.\EXPR)\VAL} \ExprTrans
      \EC{\EXPR[\VAL/x]} }

    \inferrule{\TYPE \SubType \TYPEU}{ \EC{\CAST\VAL\TYPE\TYPEU} \ExprTrans \EC{\VAL}}
    %
    % \inferrule{}{ \EC{\MATCH\ (\VAL_1, \VAL_2) \WITH\ (x_1,
    %     x_2)\to\EXPR \mid \dots} \ExprTrans \EC{\EXPR[\VAL_1,
    %     \VAL_2/x_1, x_2]} }
    % \inferrule{}{ \EC{\MATCH\ \NIL\ \WITH\ \NIL\to\EXPR \mid \dots}
    %   \ExprTrans \EC{\EXPR} }
    % \inferrule{}{ \EC{\MATCH\ (\CONS\ \VAL_1\ \VAL_2)\ \WITH\ (\CONS\
    %     x_1\ x_2)\to\EXPR \mid \dots} \ExprTrans
    %   \EC{\EXPR[\VAL_1,\VAL_2/x_1, x_2]} }
    % \inferrule{}{ \EC{\MATCH\ \NONE\ \WITH\ \NONE\to\EXPR \mid \dots}
    %   \ExprTrans \EC{\EXPR} }
    % \inferrule{}{ \EC{\MATCH\ (\SOME\ \VAL)\ \WITH\ (\SOME\ x)\to\EXPR
    %     \mid \dots} \ExprTrans \EC{\EXPR[\VAL/x]} }
    % \inferrule{}{ \EC{\MATCH\ (\LEFT\ \VAL)\ \WITH\ (\LEFT\ x)\to\EXPR
    %     \mid \dots} \ExprTrans \EC{\EXPR[\VAL/x]} }
    % \inferrule{}{ \EC{\MATCH\ (\RIGHT\ \VAL)\ \WITH\ (\RIGHT\
    %     x)\to\EXPR \mid \dots} \ExprTrans \EC{\EXPR[\VAL/x]} }
  \end{mathpar}
  \caption{Select expression reduction rules (pattern matching omitted)}
  \label{fig:expression-reduction}
\end{figure}
The internal transitions of a node are just evaluation of expressions.
\begin{mathpar}
  \inferrule[Node-Eval]
  {
    \EXPR \ExprTrans \EXPR'
  }{
    [\EC\EXPR :: \EXPRS, \ACCOUNTS] \NodeTrans{}
    [\EC{\EXPR'} :: \EXPRS, \ACCOUNTS]
  }
\end{mathpar}
The remaining transitions affect a local node in the context of the
blockchain. To this end, any local node may be selected.
\begin{mathpar}
  \inferrule[Config-System]{
    \NODE\|\BLOCKCHAIN \SystemTrans \NODE'\|\BLOCKCHAIN'
}{
    \BLOCKCHAIN[\NODE :: \overline{\NODE}] \SystemTrans
    \BLOCKCHAIN'[\NODE' :: \overline{\NODE}]
  }

  \inferrule[Config-Node]
  {\NODE \NodeTrans \NODE'}
  { {\BLOCKCHAIN[\NODE :: \overline\NODE]}
    \SystemTrans
    {\BLOCKCHAIN[\NODE' :: \overline\NODE]}}

  \inferrule[Config-Block]
  {\BLOCKCHAIN \BlockTrans \BLOCKCHAIN'}
  { \BLOCKCHAIN[{\overline\NODE}]
    \SystemTrans
    \BLOCKCHAIN'[{\overline\NODE}]}
\end{mathpar}

\begin{figure}[tp]
\begin{mathpar}
  \inferrule[Node-Inject]{
    \Angle{\PAK,\PUK} \in \ACCOUNTS \\
    \CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ) \\
    \CHECKCOU (\MANAGERS, \PUK) \\
    \CHECKPUB (\MANAGERS, \PUK') \\
    \CHECKGAS (\PUK, \PUK', \NTEZ, \MTEZ) \\
    %\OPH = \GENERATEOPH (\PUK, \PUK','', \NTEZ, \MTEZ, \TIME) \\
    \OPH = \GENERATEOPH (\OP, \TIME) \\
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ    
  }{
    { [\EC\OP :: \EXPRS, \ACCOUNTS] \|
      [\PENDING, \MANAGERS, \CONTRACTORS, \TIME] } \SystemTrans \\
   { [\EC{\OPH}  :: \EXPRS, \ACCOUNTS] \|
     [ \OPH \mapsto \Angle{\OP, \TIME, \STATUSPENDING}
     ::\PENDING,
     \UPDATECOU(\MANAGERS, \PUK, \TRUE),
     \CONTRACTORS,
     \TIME]
   }
 }

  \inferrule[Node-Reject]{
    \NEG\ \CHECKBAL (\BLOCKCHAIN.\MANAGERS, \OP.\PUK, \OP.\NTEZ, \OP.\MTEZ) \\
    }{
    { [\EC\OP :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
      } \SystemTrans
    { [\EC{\RAISE\ \ERRBAL} :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    }
  }

  \inferrule[Block-Accept]{
    \OP = \TRANSFER\NTEZ\PUK{\PUK'}\MTEZ \\
    \TIME - \hat\TIME \le 60
  }{
    { 
      [\OPH \mapsto \Angle{\OP, \hat \TIME, \STATUSPENDING}
     ::\PENDING, \MANAGERS,
      \CONTRACTORS, \TIME]}
    \BlockTrans  \\
    { 
      [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSINCLUDING\ \TIME} :: \PENDING, 
      \UPDATESUCC(\MANAGERS, \PUK, \PUK', \NTEZ, \MTEZ),
      \CONTRACTORS, \TIME +1]}
  }

    \inferrule[Block-Timeout]{
    % \OP = \TRANSFER[\STRING]\NTEZ\PUK\ADDR\MTEZ \\
    \TIME-\hat\TIME > 60
  }{ 
    {[\OPH \mapsto \Angle{\OP, \hat \TIME, \STATUSPENDING}
     ::\PENDING, \MANAGERS,
      \CONTRACTORS, \TIME]}
    \BlockTrans \\
    { 
      [\OPH \mapsto \Angle{\OP, \hat \TIME, \STATUSTIMEOUT}
     :: \PENDING, \UPDATECOU(\MANAGERS, \OP.\PUK, \FALSE),
      \CONTRACTORS, \TIME]}
  }
\end{mathpar}

  \caption{Lifecycle transitions of a transaction}
  \label{fig:lifecycle-transaction}
\end{figure}
Figure~\ref{fig:lifecycle-transaction} shows the transitions to start
and finalize a transfer transaction.
\TirName{Node-Inject} affects a local node and the blockchain. It
nondeterminstically selects a program that wants to do a transfer
operation. It
checks whether the sender of the transfer is a valid local account, whether
the balance is sufficient to pay the fee and the transferred amount,
whether there is an active transition for this sender (chkCount),
whether the receiver's address is valid, and whether the fee is
sufficient. If these conditions are fulfilled, the transition forges
an operation hash and returns it to the local node. 
The pending operation enters the pool and the sender's counter is set
to indicate an ongoing transition.

We give just one example \TirName{Node-Reject} of the numerous
transitions that cover the cases where one of the 
premises of \TirName{Node-Inject} is not fulfilled. Each of them
raises an exception that describes which condition was violated.

Acceptance or rejection of a pending operation happens on the
blockchain independent of any local node. In our model, these
transitions are nondeterministic so that acceptance can happen any
time in the next 60 cycles \TirName{Block-Accept}. Afterwards, a
pending operation can only time out \TirName{Block-Timeout}.
If the transaction is accepted, then the sender's counter is reset,
the balances of sender and receiver are adjusted (updSucc), and the
time stamp increases.

Whereas \TirName{Node-Inject} and \TirName{Block-Accept} are
particular to the transfer operation, the timeout transition applies
to all operations. It just changes the state of the operation and
resets the sender's counter, thus rolling back the transaction.

\subsection{Cast Reductions}
\label{sec:special-reductions}

\begin{figure}[tp]
\begin{mathpar}
  \inferrule[Contract-Yes]{
    \JTypeCode\CODE{\TPAIR\ \TYPE\ \TYPEU} \\
    \BLOCKCHAIN.\CONTRACTORS (\PUH) =  \Angle{\CODE, \tilde\TIME, \NTEZ', \STRING'}
    %
    }{
    { [\EC{\CAST\PUH\TPUH{\TCONTRACT\ \TYPE}}  :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    }
    \SystemTrans
    { [\EC{\PUH} :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
   }
  }

  % \inferrule[Contract-Implicit?]{
  %   \PUK \in \DOM(\BLOCKCHAIN.\MANAGERS)
  %   %
  %   }{
  %   { [\EC{\CONTRACT\ \TUNIT\ \PUK}  :: \EXPRS, \ACCOUNTS] \|
  %     \BLOCKCHAIN
  %   }
  %   \SystemTrans
  %   { [\EC{\SOME\ \PUK} :: \EXPRS, \ACCOUNTS] \|
  %     \BLOCKCHAIN
  %  }
  % }
  %
  \inferrule[Contract-No]{
    % \ADDR \in \DOM (\BLOCKCHAIN.\MANAGERS) \Rightarrow \TYPE\ne\TUNIT \\
    \BLOCKCHAIN.\CONTRACTORS (\PUH) = \Angle{\CODE, \tilde\TIME,
      \NTEZ', \STRING'} \Rightarrow {}
        \JTypeCode\CODE{\TPAIR\ \TYPE'\ \TYPEU} \wedge \TYPE \ne \TYPE' \\
    }{
    { [\EC{\CAST\PUH\TPUH{\TCONTRACT\ \TYPE}}  :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
    }
    \SystemTrans
    { [\EC{\RAISE\ \ERRPRG} :: \EXPRS, \ACCOUNTS] \|
      \BLOCKCHAIN
   }
  }
\end{mathpar}
  \caption{Cast reductions (excerpt)}
  \label{fig:cast-reductions}
\end{figure}
\begin{figure}[tp]
  \begin{mathpar}
    \inferrule[Block-Originate]{
      \Angle{\PAK,\PUK} \in \ACCOUNTS \\ \CHECKBAL (\MANAGERS, \PUK, \NTEZ, \MTEZ) \\
      \CHECKCOU (\MANAGERS, \PUK) \\
      \CHECKPRG (\CODE) \\
      \CHECKGAS (\CODE, \INIT, \NTEZ, \MTEZ)  \\
      \CHECKINIT (\CODE, \STRING) \\
      % \OPH = \GENERATEOPH(\PUK, \CODE, \STRING, \NTEZ, \MTEZ, \TIME)
      % \\
      \OPH = \GENERATEOPH(\OP, \TIME) \\
      \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ }{ [\EC\OP :: \EXPRS,
      \ACCOUNTS
      ] \| [\PENDING, \MANAGERS, \CONTRACTORS, \TIME] \to \\
      [\EC{\OPH} :: \EXPRS, \ACCOUNTS] \| [\OPH \mapsto \Angle{\OP,
        \TIME, \STATUSPENDING} ::\PENDING,
      \UPDATECOU(\MANAGERS,\PUK,\TRUE), \CONTRACTORS, \TIME] }

  \inferrule[Block-Accept]{
    \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ \\
    \PUH = \GENERATEHASH(\CODE, \TIME) \\
    \TIME-\hat\TIME  \le 60
  }{
    [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSPENDING} :: \PENDING, \MANAGERS, \CONTRACTORS, \TIME]
    \BlockTrans \\
     [\OPH \mapsto \Angle{\OP, \hat\TIME, \STATUSINCLUDING\ \TIME} :: \PENDING, \UPDATESUCC
      (\MANAGERS, \PUK, \NTEZ, \MTEZ),\\ \PUH \mapsto  \Angle{\CODE, \TIME, \NTEZ, \STRING} :: \CONTRACTORS, \TIME+1]
  }

    \inferrule[Block-Accept-Query]{
    \OP = \ORIGINATE\NTEZ\PUK\CODE\STRING\MTEZ \\
    \PENDING (\OPH) =  \Angle{\OP,
      \hat\TIME, \STATUSINCLUDING\ \tilde\TIME} \\
    \PUH = \GENERATEHASH (\CODE, \tilde\TIME)
    % \CONTRACTORS (\PUH) =  \Angle{\CODE, \tilde\TIME, \NTEZ', \STRING'}
  }{
      [\EC{\GETCONTRACT\ \OPH}  :: \EXPRS, \ACCOUNTS] \| [\PENDING, \MANAGERS, \CONTRACTORS, \TIME]
    \SystemTrans 
     [\EC{ \PUH}  :: \EXPRS, \ACCOUNTS] \|  [\PENDING, \MANAGERS, \CONTRACTORS, \TIME]
  }
\end{mathpar}
  \caption{Smart contract origination}
  \label{fig:contract-origination}
\end{figure}
\begin{figure}[tp]
  \begin{mathpar}
    \inferrule[Query-Balance-Implicit]{
      % \BLOCKCHAIN =
      % [\PENDING, \PUK \mapsto \Angle{\BAL, \COU} :: \MANAGERS,
      % \CONTRACTORS, \TIME]
      \BLOCKCHAIN.\MANAGERS (\PUK) = \Angle{\BAL,\COU}
    }{[\EC{\GETBALANCE\PUK} :: \EXPRS, \ACCOUNTS] \| \BLOCKCHAIN
      \SystemTrans\ [\EC{\BAL} ::\EXPRS, \ACCOUNTS] \| \BLOCKCHAIN}

    \inferrule[Query-Balance-Fail]{ \PUK \notin \DOM
      (\BLOCKCHAIN.\MANAGERS) \ }{[\EC{\GETBALANCE\PUK} :: \EXPRS,
      \ACCOUNTS] \| \BLOCKCHAIN \SystemTrans {[\EC{\RAISE\ \ERRPUK}
        ::\EXPRS, \ACCOUNTS] \| \BLOCKCHAIN}}
  \end{mathpar}
  \caption{Example queries}
  \label{fig:example-queries}
\end{figure}

Figure~\ref{fig:cast-reductions} contains the most interesting
example of cast reductions, from a public hash to a typed
contract. These reductions force the local node to obtain information
from the blockchain. The cast succeeds \TirName{Contract-Yes} on \PUH, if there is a
contractor for \PUH\ such that the stored code has the parameter type
expected by the cast. The cast fails \TirName{Contract-No}, if \PUH\
is invalid or if the types do not match.

\subsection{Smart Contracts}
\label{sec:smart-contracts}

The invocation of smart contracts is similar to a transfer, so we
elide the details. Figure~\ref{fig:contract-origination} contains the
transition \TirName{Block-Originate} to originate a smart contract. The basic scheme is similar
to the transfer. The preconditions for the operation are checked, but
there are extra preconditions for origination:  the program must be
well-formed and typed, the initial storage value must match its
type. The operation ends up in the pool in pending status.

Acceptance of origination is slightly different as for transfers as
shown in \TirName{Block-Accept}. We calculate the public hash \PUH\ of the
contract from the code and the current time stamp and create a new
contractor at that address.

We obtain the handle of the contract through a query, once the
contract is accepted on the blockchain in
\TirName{Block-Accept-Query}. The query's argument is the operation
hash, which is used to obtain the code and the time stamp of its
acceptance. From this information, we can re-calculate the public
hash.

\subsection{Queries}
\label{sec:queries}

We conclude with two example transitions for a simple query in
Figure~\ref{fig:example-queries}. To obtain the balance of an implicit
account \PUK, we obtain the account info from the manager and extract
the balance (\TirName{Query-Balance-Implicit}). If the account is
unknown, then we raise an exception
(\TirName{Query-Balance-Fail}). Other queries are implemented analogously.

\section{Properties}
\label{sec:properties}

\subsection{Properties of blockchain state transitions}

Running a program that starts with valid references to accounts,
operations, and contracts does not get corrupted by a transition.
\begin{proposition}
  The following property is preserved by a step on a well-formed
  configuration $ [\EXPRS, \ACCOUNTS] \| \BLOCKCHAIN$:
  \begin{itemize}
  \item for all $\OPH$ in $\EXPRS$, $\OPH \in \DOM (\BLOCKCHAIN.\PENDING)$,
  \item for all $\PUK$ in $\EXPRS$, $\PUK \in \DOM (\BLOCKCHAIN.\MANAGERS)$,
  \item for all $\PUH$ in $\EXPRS$, $\PUH \in \DOM (\BLOCKCHAIN.\CONTRACTORS)$.
  \end{itemize}
\end{proposition}

\begin{proposition}
If $[\PENDING, \MANAGERS, \CONTRACTORS, \TIME] \BlockTrans{}
[\PENDING', \MANAGERS', \CONTRACTORS', \TIME']$, then
\begin{enumerate}
\item $\TIME \le \TIME'$
\item\label{item:1} $\DOM (\PENDING) \subseteq \DOM(\PENDING')$
\item invariant for the pool: if  $\PENDING (\OPH) = \Angle{\OP,
    \hat\TIME, \STATUS}$, then $\OPH = \GENERATEOPH (\OP,
  \hat\TIME)$. 
\item for all $\OPH \in \DOM(P)$, if 
  $\PENDING (\OPH) = \Angle{\OP,  \hat\TIME, \STATUS}$, then either
  \begin{itemize}
  \item     $\PENDING'    (\OPH) = \PENDING (\OPH)$; or
  \item $\STATUS  = \STATUSPENDING$ and $\PENDING' (\OPH) =
    \Angle{\OP, \hat\TIME, \STATUSTIMEOUT}$; or
  \item  $\STATUS  = \STATUSPENDING$, $\TIME - \hat\TIME
    \le 60 $, and
    $\PENDING' (\OPH) =       \Angle{\OP, \hat\TIME, \STATUSINCLUDED\
      \TIME}$.
  \end{itemize}
 \item\label{item:2} for all $\OPH \in \DOM(P)$ and $\PENDING (\OPH) = \Angle{\OP,  \hat\TIME, \STATUSPENDING}$, 
    \begin{itemize}
    	\item  if $\MANAGERS(\OP.\PUK) =  \Angle{\BAL, \COU}$ then  
  $\COU.b = \TRUE$.
    \end{itemize}
 \item\label{item:3} $ \DOM (\MANAGERS) \subseteq \DOM (\MANAGERS')$
 \item\label{item:4} for all $\PUK \in \DOM (\MANAGERS)$\\
   if $\MANAGERS (\PUK) =
   \Angle{\BAL, \COU}$, 
   then $\MANAGERS' (\PUK) =
   \Angle{\BAL', \COU'}$ and
   \begin{itemize}
   \item if $\COU.b=\TRUE$ and $\COU'.b=\FALSE$, then $\COU.n' \in \{
     \COU.n,  \COU.n+1\}$,
   \item otherwise $\COU.n = \COU'.n$
   \item If $\COU.n = \COU'.n$, then $\BAL = \BAL'$.
   % \item if $\COU.b= \FALSE$, then $n' = n$
   % \item if $\COU.b= \TRUE$ and $b'=\TRUE$, then $n' =n$
   \end{itemize}

 \item\label{item:5} $ \DOM (\CONTRACTORS) \subseteq \DOM (\CONTRACTORS')$
   \begin{itemize}
   \item for all $\PUH \in \DOM (\CONTRACTORS)$,
     $\CONTRACTORS (\PUH).\CODE = \CONTRACTORS' (\PUH).\CODE$
   \end{itemize}
 \item\label{item:6} invariant for contractors:
   for all $\PUH \in \DOM (\CONTRACTORS)$,
   $\CONTRACTORS (\PUH) = \Angle{ \CODE, \tilde\TIME, \BAL, \STORAGE}$
   implies that $\PUH = \GENERATEHASH (\CODE, \tilde\TIME)$.
\end{enumerate}
\end{proposition}
Establishing these properties relies on the preimage resistance of the
various hash functions used to calculate operation hashes and public
hashes. Items~\ref{item:1}--\ref{item:2} describe an invariant and the lifecycle of
operations. Items~\ref{item:3} and \ref{item:4} describe the lifecycle
of a transfer and items~\ref{item:5} and~\ref{item:6} describe
invariants for contractors. The invariants establish the
self-verifying property common of blockchain entities.

@@

To describe contracts we associate each public hash with a pair of
types, the parameter type and the storage type.
\begin{align*}
  \Delta & ::= \EmptyEnv \mid \PUH: \TPAIR\ \TYPE\ \TYPEU, \Delta
\end{align*}

@@

\newpage{}
\appendix

\section{Unused}
\label{sec:unused}

\subsection{An Implicit Account Transfer}
